---
title: "Sampling vs. Jittering: Tapping variability"
author: Jelle van der Werff
date: 02/09/2022
output:
    html_document:
        code_folding: hide
        toc: true
---

# Introduction
Here, we aim to answer the question: Does the underlying regularity in jittering sequences lead to more regular
tapping?

The DV is:

* $\textrm{SD}_\textrm{ITI}$, which is the variability in the intervals between response taps in a specific trial.


Though $\textrm{SD}_\textrm{async}$ is also an often-used variable (see Repp & Su, 2013), it is not used here, as
it was considered not so useful in contexts where the stimulus is as anisochronous as it is here.

The independent variables used here are:

* Condition (Sampling or jittering)
* Sequence length (25 or 35)
* Sequence tempo (400 ms or 500 ms)
* Amount of variance (less or more)



```{r preliminaries, warning = FALSE, message = FALSE}
############# PACKAGES AND OTHER PRELIMINARIES ############


# for easy loading of packages
if (!require("pacman")) install.packages("pacman")

# load packages
p_load(
  "plyr", "dplyr", "nlme", "here", "ggplot2", "reshape2", "ggsci", "knitr", "kableExtra", "scico", "cowplot",
  "effectsize", "sjPlot", "emmeans"
)
here <- here::here

# for relative file paths (e.g. here('Input', 'file.csv'))
here::i_am(file.path("scripts", "tapping", "variability", "variability.rmd"))

# Load data
ITIs <- readRDS(here("data", "tapping_clean", "tapping_intervals.rds"))

# load custom ggplot theme
source(here("scripts", "custom_ggplot_theme.R"))

# Labels for plots
labels <- as_labeller(c(
  "400" = "Faster tempo\n(400 ms)",
  "500" = "Slower tempo\n(500 ms)",
  "Sampling" = "Sampling",
  "Jittering" = "Jittering",
  "0.25" = "More variance",
  "0.2" = "Less variance",
  "25" = "Shorter (25 events)",
  "35" = "Longer (35 events)"
))

# Knitr settings
knitr::opts_chunk$set(cache = TRUE, out.width = "100%", warning = FALSE, message = FALSE)
```

# Descriptives

## Table
```{r desc-table}
ITIs$cov_iti <- (ITIs$resp_sd / ITIs$resp_tempo) * 100
ITIs$cov_ioi <- (ITIs$stim_sd / ITIs$stim_tempo) * 100

tapping_variability_descriptives <- ddply(ITIs, .(tempo, variance, length, condition), summarize,
  n = length(resp_iti),
  ioi_mean = mean(stim_ioi, na.rm = TRUE),
  iti_mean = mean(resp_iti, na.rm = TRUE),
  ioi_sd = sd(stim_ioi, na.rm = TRUE),
  iti_sd = sd(resp_iti, na.rm = TRUE)
)



kable(tapping_variability_descriptives) %>% kable_styling(bootstrap_options = c("striped", "hover"))
```


# Summarize SDs

```{r summarize-sds}
ITIs.sd <- ddply(ITIs, .(tempo, variance, length, condition, pp_id, age, handedness, gender, sequence_id), summarize,
  iti_sd = sd(resp_iti, na.rm = TRUE)
)
# remove missing values
ITIs.sd <- ITIs.sd[!is.na(ITIs.sd$iti_sd), ]


ITIs.sd$pp_id <- as.factor(ITIs.sd$pp_id)
ITIs.sd$length <- as.factor(ITIs.sd$length)
ITIs.sd$variance <- as.factor(ITIs.sd$variance)
ITIs.sd$tempo <- as.factor(ITIs.sd$tempo)
ITIs.sd$condition <- as.factor(ITIs.sd$condition)
ITIs.sd$handedness <- as.factor(ITIs.sd$handedness)
ITIs.sd$gender <- as.factor(ITIs.sd$gender)
```

# LME
## Check normality

```{r check-normality}
ggplot(ITIs.sd, aes(
  x = iti_sd, colour = factor(condition, levels = c("Sampling", "Jittering")),
  fill = factor(condition, levels = c("Sampling", "Jittering"))
)) +
  geom_histogram() +
  facet_grid(tempo + variance ~ factor(condition, levels = c("Sampling", "Jittering")) + length, labeller = labels) +
  labs(x = expression(italic(SD)[ITI]), y = "Count") +
  theme_jelle() +
  scale_fill_manual(values = c("Sampling" = colours$sampling, "Jittering" = colours$jittering)) +
  scale_color_manual(values = c("Sampling" = colours$sampling, "Jittering" = colours$jittering)) +
  theme(legend.position = "bottom", legend.justification = "center")

ggsave(here("plots", "tapping", "assumptions", "sd_normality.pdf"))
ggsave(here("plots", "tapping", "assumptions", "sd_normality.png"), dpi = 600)
```

## Modelling
```{r sd-lme}
# start with baseline
baseline <- gls(iti_sd ~ 1, method = "ML", data = ITIs.sd, na.action = na.omit)

# add in random intercept
random <- lme(iti_sd ~ 1,
  random = ~ 1 | pp_id,
  method = "ML",
  data = ITIs.sd,
  control = lmeControl(opt = "optim"),
  na.action = na.omit
)
anova(baseline, random) # significant

# effect of tempo
tempo <- update(random, . ~ . + tempo)
anova(tempo, random) # significant

# effect of variance
variance <- update(tempo, . ~ . + variance)
anova(variance, tempo) # significant

# effect of condition
condition <- update(variance, . ~ . + condition)
anova(condition, variance) # significant

# effect of length
length <- update(condition, . ~ . + length)
anova(length, condition) # significant

# effect of age
age <- update(length, . ~ . + age)
anova(age, length) # not significant

# effect of handedness
handedness <- update(length, . ~ . + handedness)
anova(handedness, length) # not significant

# effect of gender
gender <- update(length, . ~ . + gender)
anova(gender, length) # not significant

# interaction effect of condition x tempo
condition.tempo <- update(length, . ~ . + condition:tempo)
anova(condition.tempo, length) # not significant

# interaction effect of condition x variance
condition.variance <- update(length, . ~ . + condition:variance)
anova(condition.variance, length) # significant

# interaction effect of condition x length
condition.length <- update(condition.variance, . ~ . + condition:length)
anova(condition.length, condition.variance) # significant


sd_model <- condition.length
summary(sd_model)
effectsize(sd_model)
```


## Plot model

```{r plot-model-interaction-length}
plot_lme <- plot_model(sd_model, type = "int")

plot_build <- ggplot_build(plot_lme[[2]])

ggplot(plot_build$data[[2]], aes(x = factor(group), colour = factor(c(2, 2, 1, 1)))) +
  geom_pointrange(aes(y = y, ymin = ymin, ymax = ymax)) +
  geom_line(aes(x = factor(group), y = y, group = factor(c(2, 2, 1, 1)))) +
  labs(x = "Length", colour = "Condition", y = expression(italic(SD)[ITI])) +
  theme_jelle() +
  scale_x_discrete(labels = c("25 events", "35 events")) +
  scale_colour_manual(values = c("1" = colours$sampling, "2" = colours$jittering)) +
  theme(legend.position = "none")
ggsave(here("plots", "tapping", "sd_interaction_conditionlength.pdf"), height = 4, width = 4)
```



```{r plot-model-interaction-variance}
plot_lme <- plot_model(sd_model, type = "int")

plot_build <- ggplot_build(plot_lme[[1]])

ggplot(plot_build$data[[2]], aes(x = factor(c(2, 2, 1, 1)), colour = factor(group))) +
  geom_pointrange(aes(y = y, ymin = ymin, ymax = ymax)) +
  geom_line(aes(x = factor(c(2, 2, 1, 1)), y = y, group = factor(group))) +
  labs(x = "Variance", colour = "Condition", y = expression(italic(SD)[ITI])) +
  scale_x_discrete(labels = c("Less variance", "More variance")) +
  scale_color_scico_d(palette = "roma", labels = c("Sampling", "Jittering"))
ggsave(here("plots", "tapping", "sd_interaction_conditionvariance.eps"), scale = 0.4, height = 6.5, width = 10)
```

```{r forestplot}
plot_model(sd_model, type = "std", axis.labels = rev(c("Tempo [500 ms]", "Variance [More]", "Condition [Jittering]", "Length [Longer]", "Interaction [Variance x Condition]", "Interaction [Length x Condition]"))) +
  labs(title = "Standardized effect sizes for tapping variability", subtitle = expression("Mixed-effects model for" ~ SD[ITI])) +
  theme_jelle() +
  theme(legend.position = "none") +
  theme(plot.title = element_text(hjust = 0)) +
  ylim(-0.5, 1) +
  geom_hline(yintercept = c(-0.5, 0, 0.5, 1), linetype = "dashed", color = "#cbcbcb")

ggsave(here("plots", "tapping", "sd_forestplot.png"), scale = 0.75)
ggsave(here("plots", "tapping", "sd_forestplot.eps"), scale = 0.75)
```

# Marginal means

```{r emmeans-allconditions}
mmeans <- data.frame(emmeans(sd_model, ~ condition + tempo + variance + length))
mmeans$condition <- factor(mmeans$condition, levels = c("Sampling", "Jittering"))

mmeans$condition_otherfactors_code <- paste(mmeans$tempo, mmeans$variance, mmeans$length, sep = "_")
mmeans$condition_otherfactors <- factor(mmeans$condition_otherfactors_code)

ggplot(
  mmeans,
  aes(
    y = condition_otherfactors_code,
    x = emmean,
    xmin = lower.CL,
    xmax = upper.CL,
    colour = condition,
    group = condition_otherfactors_code
  )
) +
  geom_pointrange(show.legend = FALSE) +
  theme_jelle() +
  labs(
    x = "", y = ""
  ) +
  scale_colour_manual(values = c("Sampling" = colours$sampling, "Jittering" = colours$jittering)) +
  guides(colour = guide_legend(reverse = TRUE)) +
  theme(axis.text.y = element_blank())

ggsave(here("plots", "tapping", "sd_marginalmeans.pdf"), width = 3, height = 4)
```