---
title: "Sampling vs Jittering: Tapping tempo"
author: Jelle van der Werff
date: 06/09/2022
output:
    html_document:
        code_folding: hide
        toc: true
---


# Introduction

The question here is, can people use the underlying regularity in jittering sequences to tap
closer to the sequence tempo? From the mathematical model, it follows that for jittering sequences the
tempo is more easily estimated.

Used measures:

* Tempo error: $\textrm{error}_i = |\textrm{ITI}_i - \textrm{tempo}|$
* The inter-tap interval (ITI; the interval between two consecutive taps)

```{r load-packages, message = FALSE, warning = FALSE}
# for easy loading of packages
if (!require("pacman")) install.packages("pacman")

# load packages
p_load(
  "ggplot2", "nlme", "effectsize", "knitr", "kableExtra", "plyr", "dplyr", "reshape2", "ggsci", "here",
  "sjPlot", "scico", "cowplot", "emmeans", "ggdist"
)

# use here from here instead of from plyr
here <- here::here

# for relative file paths (e.g. here('Input', 'file.csv'))
here::i_am(file.path("scripts", "tapping", "tempo", "tempo.rmd"))

# get custom ggplot theme
source(here("scripts", "custom_ggplot_theme.R"))

# Knitr settings
knitr::opts_chunk$set(cache = TRUE, out.width = "100%", warning = FALSE, message = FALSE)
```

```{r load-data}
# Load data
ITIs <- readRDS(here("data", "tapping_clean", "tapping_intervals.rds"))

# Make e_tempo variable
ITIs$resp_tempo_error <- ITIs$resp_iti - as.integer(as.character(ITIs$stim_tempo)) # we use the theoretical tempo
ITIs$resp_tempo_error_abs <- abs(ITIs$resp_tempo_error)

# Do log transform
ITIs$resp_tempo_error_abs_log <- log(ITIs$resp_tempo_error_abs)

# Remove missing values
ITIs <- ITIs[(!is.na(ITIs$resp_tempo_error_abs_log)) & (!is.infinite(ITIs$resp_tempo_error_abs_log)), ]

# Labels for plots
labels <- as_labeller(c(
  "400" = "Faster tempo\n(400 ms) ",
  "500" = "Slower tempo\n(500 ms)",
  "Sampling" = "Sampling",
  "Jittering" = "Jittering",
  "0.25" = "More variance",
  "0.2" = "Less variance",
  "25" = "Shorter (25 events)",
  "35" = "Longer (35 events)"
))
```


# Descriptives

## Tables

Here's the tapping and sequence tempo split out only by the condition and the intended tempo:

```{r tempo-table}
tempo <- ddply(ITIs, .(condition, tempo), summarize,
  n = length(resp_iti),
  mean_stim = mean(stim_ioi, na.rm = TRUE),
  mean_tap = mean(resp_iti, na.rm = TRUE),
  sd_tap = sd(resp_iti, na.rm = TRUE)
)

kable(tempo) %>% kableExtra::kable_styling(bootstrap_options = "striped")
```

Here it is split out by tempo, amount of variance, length, and condition:

```{r tempo-table-2}
tempo <- ddply(ITIs, .(tempo, variance, length, condition), summarize,
  n = length(!is.na(resp_iti)),
  mean_stim = mean(stim_ioi, na.rm = TRUE),
  mean_tap = mean(resp_iti, na.rm = TRUE),
  sd_tap = sd(resp_iti, na.rm = TRUE)
)

kable(tempo) %>% kableExtra::kable_styling(bootstrap_options = "striped")
```


# Tapping

```{r tapping-error}
tempo.error <- ddply(ITIs, .(tempo, variance, length, condition, interval_i), summarize,
  stimulus = mean(stim_ioi, na.rm = TRUE) - tempo,
  response = mean(resp_iti, na.rm = TRUE) - tempo,
)
```



# Modelling

## Check normality
```{r check-normality}
ggplot(ITIs, aes(
  x = resp_tempo_error_abs_log,
  colour = factor(condition, levels = c("Sampling", "Jittering")),
  fill = factor(condition, levels = c("Sampling", "Jittering"))
)) +
  geom_histogram() +
  facet_grid(tempo + variance ~ factor(condition, levels = c("Sampling", "Jittering")) + length, labeller = labels) +
  labs(x = expression(log ~ italic(e)[tempo]), y = "Count") +
  theme_jelle() +
  scale_fill_manual(values = c("Jittering" = colours$jittering, "Sampling" = colours$sampling)) +
  scale_color_manual(values = c("Jittering" = colours$jittering, "Sampling" = colours$sampling)) +
  theme(legend.position = "bottom", legend.justification = "center")

ggsave(here("plots", "tapping", "assumptions", "e_tempo_normality.pdf"))
ggsave(here("plots", "tapping", "assumptions", "e_tempo_normality.png"))
```


* __A__: Sampling, length = 25
* __B__: Sampling, length = 35
* __C__: Jittering, length = 25
* __D__: Jittering, length = 35


We're not doing K-S/Shapiro-Wilk because our sample is too big.



## Model building
```{r model-tempo-error}
# start with baseline
baseline <- gls(resp_tempo_error_abs_log ~ 1, method = "ML", data = ITIs, na.action = na.omit)

# add in random intercept
random <- lme(resp_tempo_error_abs_log ~ 1,
  random = ~ 1 | pp_id,
  method = "ML",
  data = ITIs,
  control = lmeControl(opt = "optim"),
  na.action = na.omit
)
anova(baseline, random) # significant

# effect of tempo
tempo <- update(random, . ~ . + tempo)
anova(tempo, random) # significant

# effect of variance
variance <- update(tempo, . ~ . + variance)
anova(variance, tempo) # significant

# effect of condition
condition <- update(variance, . ~ . + condition)
anova(condition, variance) # significant

# effect of length
length <- update(condition, . ~ . + length)
anova(length, condition) # significant

# effect of age
age <- update(length, . ~ . + age)
anova(age, length) # not significant

# effect of handedness
handedness <- update(length, . ~ . + handedness)
anova(handedness, length) # not significant

# effect of gender
gender <- update(length, . ~ . + gender)
anova(gender, length) # not significant

# interaction effect of condition x tempo
condition.tempo <- update(length, . ~ . + condition:tempo)
anova(condition.tempo, length) # significant

# interaction effect of condition x variance
condition.variance <- update(condition.tempo, . ~ . + condition:variance)
anova(condition.variance, condition.tempo) # not significant

# interaction effect of condition x length
condition.length <- update(condition.tempo, . ~ . + condition:length)
anova(condition.tempo, condition.length) # significant


error_model <- condition.length
summary(error_model)
effectsize(error_model)
```

## Plot model

```{r plot-lme}
plot_lme <- plot_model(error_model, type = "pred", terms = "condition")

plot_build <- ggplot_build(plot_lme)

ggplot(data = plot_build$data[[2]], aes(x = factor(x), colour = factor(x))) +
  geom_pointrange(aes(ymin = ymin, ymax = ymax, y = y)) +
  labs(x = "Condition", colour = "Condition", y = expression(log ~ italic(e)[tempo])) +
  scale_x_discrete(labels = c("Sampling", "Jittering")) +
  theme_jelle() +
  theme(legend.position = "none", axis.title.x = element_blank())
ggsave(here("plots", "tapping", "e_tempo_predicted_bycondition.pdf"), scale = 0.75, height = 6.5, width = 7)
```

```{r plot-interaction-length}
plot_lme <- plot_model(error_model, type = "int")

plot_build <- ggplot_build(plot_lme[[2]])

ggplot(plot_build$data[[2]], aes(colour = factor(c(2, 2, 1, 1)), x = factor(group))) +
  geom_pointrange(aes(y = y, ymin = ymin, ymax = ymax)) +
  geom_line(aes(x = factor(group), y = y, group = factor(c(2, 2, 1, 1)))) +
  labs(x = "Length", colour = "Condition", y = expression(log ~ italic(e)[tempo])) +
  theme_jelle() +
  scale_x_discrete(labels = c("25 events", "35 events")) +
  scale_colour_manual(values = c("1" = colours$sampling, "2" = colours$jittering)) +
  theme(legend.position = "none")

ggsave(here("plots", "tapping", "e_tempo_interaction_conditionlength.pdf"), height = 4, width = 4)
```


```{r plot-interaction-tempo}
plot_lme <- plot_model(error_model, type = "int")

plot_build <- ggplot_build(plot_lme[[1]])

ggplot(plot_build$data[[2]], aes(x = factor(c(1, 1, 2, 2)), colour = factor(group))) +
  geom_pointrange(aes(y = y, ymin = ymin, ymax = ymax), alpha = 0.65) +
  geom_line(aes(x = factor(c(1, 1, 2, 2)), y = y, group = factor(group)), alpha = 0.65) +
  labs(x = "Tempo", colour = "Condition", y = expression(log ~ italic(e)[tempo])) +
  scale_x_discrete(labels = c("Faster tempo\n(400 ms)", "Slower tempo\n(500 ms)")) +
  theme_jelle() +
  scale_color_scico_d(palette = "roma", labels = c("Sampling", "Jittering")) +
  theme(legend.position = "none")

ggsave(here("plots", "tapping", "e_tempo_interaction_conditiontempo.pdf"), height = 4, width = 4)
```



```{r forest-plot}
plot_model(error_model, type = "std", axis.labels = rev(c("Tempo [500 ms]", "Variance [More]", "Condition [Jittering]", "Length [Longer]", "Interaction [Tempo x Condition]", "Interaction [Length x Condition]"))) +
  labs(title = "Standardized effect sizes for tempo estimation error", subtitle = expression("Mixed-effects model for" ~ e[tempo])) +
  theme_jelle() +
  ylim(-0.25, 0.5) +
  geom_hline(yintercept = c(-0.25, 0, 0.25, 0.5), linetype = "dashed", colour = "#cbcbcb") +
  theme(legend.position = "none") +
  theme(plot.title = element_text(hjust = 0))

ggsave(here("plots", "tapping", "e_tempo_forestplot.png"), scale = 0.75)
ggsave(here("plots", "tapping", "e_tempo_forestplot.eps"), scale = 0.75)
```

```{r etempo-by-intervalnumber}
ggplot(ITIs[ITIs$interval_i %in% 3:20, ], aes(x = factor(interval_i), y = resp_tempo_error_abs, colour = condition, group = condition)) +
  stat_summary(geom = "pointrange") +
  geom_smooth(method = "lm") +
  labs(
    title = expression(bold(e[tempo]) ~ bold("by interval number")),
    x = expression(ITI[i]), y = expression(Absolute ~ e[tempo]), subtitle = "Collapsed across lengths, tempi, and degrees of variance"
  ) +
  theme_jelle() +
  theme(plot.subtitle = element_text(face = "plain"))
ggsave(here("plots", "tapping", "e_tempo_byintervalnumber.pdf"), scale = 0.75)
ggsave(here("plots", "tapping", "e_tempo_byintervalnumber.png"), scale = 0.75)
```



# Marginal means

```{r emmeans-plot-all}
mmeans <- data.frame(emmeans(error_model, ~ condition + tempo + variance + length))
mmeans$condition <- factor(mmeans$condition, levels = c("Sampling", "Jittering"))

mmeans$condition_otherfactors_code <- paste(mmeans$tempo, mmeans$variance, mmeans$length, sep = "_")
mmeans$condition_otherfactors <- factor(mmeans$condition_otherfactors_code)

ggplot(mmeans, aes(y = condition_otherfactors_code, x = emmean, xmin = lower.CL, xmax = upper.CL, colour = condition, group = condition_otherfactors_code)) +
  geom_pointrange(show.legend = FALSE) +
  theme_jelle() +
  labs(
    x = "", y = ""
  ) +
  scale_colour_manual(values = c("Sampling" = colours$sampling, "Jittering" = colours$jittering)) +
  guides(colour = guide_legend(reverse = TRUE)) +
  theme(axis.text.y = element_blank(
  ))

ggsave(here("plots", "tapping", "e_tempo_marginalmeans.pdf"), width = 3, height = 4)
```