---
title: "Sampling vs. Jittering: Continuation phase analysis"
author: Jelle van der Werff
date: 02/09/2022
output:
    html_document:
        code_folding: hide
        toc: true
---

# Load packages

```{r preliminaries, warning = FALSE, message = FALSE}
# for easy loading of packages
if (!require("pacman")) install.packages("pacman")

# load packages
p_load("here", "ggplot2", "nlme", "plyr", "dplyr", "sjPlot", "scico", "effectsize", "cowplot")

# for relative file paths (e.g. here('Input', 'file.csv'))
here <- here::here
here::i_am(file.path("scripts", "tapping", "continuation", "continuation.rmd"))

## Load data
ITIs <- readRDS(here("data", "tapping_clean", "tapping_synchronization_and_continuation_ITIs.rds"))


# Labels for plots
labels <- as_labeller(c(
        "400" = "Faster tempo\n(400 ms)",
        "500" = "Slower tempo\n(500 ms)",
        "sampling" = "Sampling",
        "jittering" = "Jittering",
        "0.25" = "More variance",
        "0.2" = "Less variance",
        "25" = "Shorter (25 events)",
        "35" = "Longer (35 events)"
))

# get custom ggplot theme
source(here("scripts", "custom_ggplot_theme.R"))

# Change condition into factor with named levels
ITIs$condition <- as.factor(ITIs$condition)
levels(ITIs$condition) <- c("Jittering", "Sampling")

# Knitr settings
knitr::opts_chunk$set(cache = TRUE, out.width = "100%", warning = FALSE, message = FALSE)
```

# Plot

We only want the 4 intervals that precede interval no. 24/34, and the 4 intervals from 26-29 (or 36-39).

```{r select-intervals}
ITIs <- ITIs[(ITIs$length == 25 & ITIs$interval_i %in% c(20:23, 26:29)) | (ITIs$length == 35 & ITIs$interval_i %in% c(30:33, 36:39)), ]
```


```{r simple-plot}
ITIs.forplot <- ITIs[(ITIs$length == 25 & ITIs$interval_i %in% 20:29) | (ITIs$length == 35 & ITIs$interval_i %in% 30:39), ]

ITIs.forplot$condition <- factor(ITIs.forplot$condition, levels = c("Sampling", "Jittering"))

ggplot(ITIs.forplot, aes(x = factor(interval_i), y = resp_iti, colour = condition, group = condition)) +
        stat_summary(geom = "pointrange") +
        stat_summary(geom = "line") +
        facet_grid(tempo + variance ~ length, scales = "free_x", labeller = labels) +
        labs(x = "Interval number", y = "Inter-tap interval (milliseconds)", colour = "Condition") +
        theme_jelle() +
        theme(legend.position = "bottom", legend.justification = "center") +
        scale_colour_manual(values = c("Sampling" = colours$sampling, "Jittering" = colours$jittering)) +
        geom_vline(data = ITIs.forplot[ITIs.forplot$length == 25, ], aes(xintercept = 4.5), linetype = "dashed") +
        geom_vline(data = ITIs.forplot[ITIs.forplot$length == 35, ], aes(xintercept = 4.5), linetype = "dashed")

ggsave(here("plots", "tapping", "continuation_itis.pdf"), width = 12, height = 8)
```


# Is tapping closer to the stimulus tempo after the final isochronous event?

We take into account only the final 4 intervals before the isochronous event for the synchronization phase.
We do this because there is motor/tempo drift towards the end of the sequence.
and compare them to the 4 following the isochronous event for the continuation phase.

```{r calculate-abs-error}
ITIs$e_tempo <- abs(ITIs$resp_iti - ITIs$stim_tempo)
ITIs$e_tempo_log <- log(ITIs$e_tempo) # do log transform
```


```{r e_tempo_plot}
ggplot(ITIs, aes(x = factor(interval_i), y = e_tempo, colour = condition, group = condition)) +
        stat_summary(geom = "pointrange") +
        stat_summary(geom = "line") +
        facet_grid(variance + tempo ~ length, scales = "free_x", labeller = labels)
```





```{r model-e_tempo}
# start with baseline
baseline <- gls(e_tempo_log ~ 1, method = "ML", data = ITIs, na.action = na.omit)

# add in random intercept
random <- lme(e_tempo_log ~ 1,
        random = ~ 1 | pp_id,
        method = "ML",
        data = ITIs,
        control = lmeControl(opt = "optim"),
        na.action = na.omit
)
anova(baseline, random) # significant

# effect of tempo
tempo <- update(random, . ~ . + tempo)
anova(tempo, random) # significant

# effect of variance
variance <- update(tempo, . ~ . + variance)
anova(variance, tempo) # significant

# effect of experimental phase (i.e. synchronization or continuation)
phase <- update(variance, . ~ . + exp_phase)
anova(phase, variance) # not significant

e_tempo_model <- phase
summary(e_tempo_model)
effectsize(e_tempo_model)
```





```{r plot-model-etempo}
plot_model(e_tempo_model, colors = "black", show.legend = FALSE, type = "std", dot.size = 0.75) +
        scale_x_discrete(labels = rev(c("Tempo [500]", "Variance [More]", "Trial phase [Sync.]"))) +
        labs(title = NULL, y = expression(Standardized ~ beta ~ group("(", log ~ e[tempo], ")"))) +
        theme_jelle() +
        scale_colour_manual(values = c("black", "black")) +
        theme(
                panel.grid.major.x = ggplot2::element_line(
                        color = "#cbcbcb", linetype = "dashed"
                ),
                legend.position = "none",
                axis.title.y = element_text(family = "Helvetica", face = "bold")
        ) +
        ylim(-0.5, 0.5)

ggsave(here("plots", "tapping", "continuation_estimates_e_tempo.pdf"), width = 4, height = 3)
```




# Is there more consistent tapping after the final isochronous event?

```{r calculate-sd}
ITIs.sd <- ddply(ITIs, .(pp_id, condition, variance, tempo, length, sequence_id, exp_phase), summarize, sd = sd(resp_iti, na.rm = TRUE))
```


```{r model-sd}
# start with baseline
baseline <- gls(sd ~ 1, method = "ML", data = ITIs.sd, na.action = na.omit)
# add in random intercept
random <- lme(sd ~ 1,
        random = ~ 1 | pp_id,
        method = "ML",
        data = ITIs.sd,
        control = lmeControl(opt = "optim"),
        na.action = na.omit
)
anova(baseline, random) # significant

# effect of tempo
tempo <- update(random, . ~ . + tempo)
anova(tempo, random) # significant

# effect of variance
variance <- update(tempo, . ~ . + variance)
anova(variance, tempo) # significant

# effect of experimental phase (synchronization vs. continuation)
exp_phase <- update(variance, . ~ . + exp_phase)
anova(exp_phase, variance) # significant

# effect of condition (Sampling vs. jittering)
condition <- update(exp_phase, . ~ . + condition)
anova(condition, exp_phase) # significant

# effect of length
length <- update(condition, . ~ . + length)
anova(condition, length) # not significant

# interaction effect of condition and phase
condition.phase <- update(condition, . ~ . + condition:exp_phase)
anova(condition.phase, condition) # not significant

sd_model <- condition.phase
summary(sd_model)
effectsize(sd_model)
```

```{r sd-forest-plot}
plot_model(sd_model, colors = "black", show.legend = FALSE, type = "std", dot.size = 0.75) +
        scale_x_discrete(labels = rev(c("Tempo [500]", "Variance [More]", "Trial phase [Sync.]", "Cond. [Sampling]", expression(Cond. %*% phase)))) +
        labs(title = NULL, y = expression(Standardized ~ beta ~ group("(", SD[ITI], ")"))) +
        theme_jelle() +
        scale_colour_manual(values = c("black", "black")) +
        theme(legend.position = "none") +
        theme(panel.grid.major.x = ggplot2::element_line(color = "#cbcbcb", linetype = "dashed")) +
        ylim(-0.5, 0.5)

ggsave(here("plots", "tapping", "continuation_estimates_sd.pdf"), width = 4, height = 3)
```
